---
title: "AARMP Descriptive"
author: "Dakota Murray"
output: html_notebook
---

```{r}
library(dplyr)
aarmp <- read.csv("/Users/dakotamurray/Dropbox/AcademicAnalytics/Data/Derived/Processed/AARMP/aarmp_processed.csv")
aadata <- read.csv("/Users/dakotamurray/Dropbox/AcademicAnalytics/Data/Derived/Processed/AA/aadata2016_processed.csv") %>% 
  group_by(AA.PersonId, AA.ProgramId) %>%
  slice(1)

rmp <- read.csv("/Users/dakotamurray/Dropbox/AcademicAnalytics/Data/Derived/Processed/RMP/rmp2017_enriched.csv")
```


## Descriptive statistics of the data
### Number of records in the AA2016 dataset
```{r}
dim(aadata)[1]
```

### Number of records (including duplicate disciplinary records) for tenure and tenure-track professors in the AA2016 dataset
```{r}
tt_profs <- aadata %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor"))
```


### Number of distinct tenure and tenure track professors in the AA2016 dataset
```{r}
dim(aadata %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor")) %>%
  group_by(AA.PersonId) %>%
  filter(row_number() == 1))[1]
```

### Number of institutions represented by tenure and tenure track professors in AA2016
```{r}
sum(!is.na(unique(tt_profs$AA.InstitutionName)))
```

### Number of level 1 disciplinary classifications in the AA2016 dataset
```{r}
length(unique(tt_profs$AA.L1Id))
```

### Number of individuals with at duplicate records in AA2016
```{r}
aa_filtered <- aadata %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor")) %>%
  group_by(AA.PersonId) %>%
  summarize(has_duplicate = n() > 1)

sum(aa_filtered$has_duplicate)
```

### Number of individuals who have duplicate professors with distinct NSF classifications
```{r}
tt_profs <- aadata %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor")) %>%
  group_by(AA.PersonId) %>%
  summarize(
    n_unique_nsf = length(unique(`OUR.Level.4`))
  )

sum(tt_profs$n_unique_nsf > 1)
```


## Descrictive statistics detailing the matching process

## Percentages of AA dataset that was matched
```{r}
tt_profs_total <- aadata %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor")) %>%
  group_by(AA.PersonId) %>%
  filter(row_number() == 1)

tt_profs_matched <- aarmp %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor")) %>%
  group_by(AA_id) %>%
  filter(row_number() == 1)


unique_total <- dim(tt_profs_total)[1]
unique_matched <- dim(tt_profs_matched)[1]
print(paste("Total number of AA individuals: ", unique_total))
print(paste("Number of AA individuals matched to RMP records: ", unique_matched))
print(paste("Percentage of AA individuals matched to RMP records: ", round(unique_matched / unique_total, 4)))
```

What percentage of RMP dataset was matched?
```{r}
rmp_profs_total <- rmp %>%
  group_by(RMP.ProfessorId) %>%
  filter(row_number() == 1)

tt_profs_matched <- aarmp_matched %>%
  filter(AA.RankType %in% c("Assistant Professor", "Associate Professor", "Professor")) %>%
  group_by(AA_id) %>%
  filter(row_number() == 1)

unique_total_all <- dim(rmp_profs_total)[1]
unique_total <- dim(rmp_profs_total %>% filter(!is.na(OUR.uniid)))[1]
unique_matched <- dim(tt_profs_matched)[1]

print(paste("Total number of individuals in Full Dataset (All Universities): ", unique_total_all))
print(paste("Number of RMP individuals matched to AA records: ", unique_matched))
print(paste("Percentage of matched to RMP records for All Universities: ", round(unique_matched / unique_total_all * 100, 4)))
print(paste("Percentage of matched to RMP records for AA Universities: ", round(unique_matched / unique_total * 100, 4)))
```

### Number of individuals in final matched dataset
```{r}
dim(aarmp)[1]
```

