---
title: "AARMP: Analysis"
author: "Dakota Murray"
output: html_notebook
---

# Load necessary packages and data
```{r}
library(tidyverse)
library(stargazer)
library(grid)
library(gridExtra)
# Load the data
ready_data <- readr::read_csv("~/Desktop/AARMP_2017/processed_files/aarmp_ready_data.csv", 
    col_types = cols(X1 = col_skip()))

# to make sure it loaded correctly, print the dimensions
dim(ready_data)
```


Perform data processing. 
```{r}
# Filter to include only those for whom a gender could be identified
ready_data <- ready_data[ready_data$gender %in% c("M", "F"), ]

# Remove NA values...
ready_data <- na.omit(ready_data) %>%
  # convert age into decades
  mutate(scientific_age = scientific_age / 10)

dim(ready_data)
```

```{r}
p_point <- ready_data %>%
  ggplot(aes(y = norm_all_pubcount, x = overall)) +
  geom_point() +
  facet_wrap(~discipline) +
  #geom_smooth(method = "loess", color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  theme_minimal()

p_point
```


```{r}
p_box <- ready_data %>%
  mutate(all_output = factor(all_output, levels = c("None", "Moderate", "High"))) %>%
  ggplot(aes(x = all_output, y = overall)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.02, width = 0.2) +
  theme_minimal() +
  facet_wrap(~discipline)

p_box
```

We can attmept to dive deeper by looking at the overall teaching quality of only those with output, consdiering them by percentile bins
```{r}


long_data <- ready_data %>%
  gather(metric, value, norm_citations, norm_all_pubcount, norm_grantcount, norm_awardcount) %>%
  mutate(
    metric = factor(metric, labels = c("Publications", "Awards", "Citations", "Grants"))
  ) 

no_count <- long_data %>%
  filter(value == 0) %>%
  group_by(metric) %>%
  summarize(
    med = median(overall)
  )

long_data %>%
  filter(value > 0) %>%
  group_by(metric) %>%
  mutate(decile_rank = ntile(value, 10)) %>%
  select(overall, metric, value, decile_rank, discipline) %>%
  ggplot(aes(x = as.factor(decile_rank), y = overall)) +
    geom_boxplot() +
    stat_summary(fun.y=median, geom="line", aes(group=1), color = "red")  +
    stat_summary(fun.y=median, geom="point", aes(group=1), color = "red")  + 
    facet_grid(~metric) +
    geom_hline(data = no_count, aes(yintercept = med), color = "black", alpha = 0.6, size = 1) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 12)
    ) +
    labs(x = "Decile ranks",
         y = "Overall Teaching Quality")

ggsave("~/Desktop/aarmp_quantile_boxplots.pdf")
```

```{r}
long_data %>%
  filter(value > 0) %>%
  #filter(metric %in% c("Citations", "Publications", "Grants")) %>%
  group_by(metric) %>%
  mutate(decile_rank = as.factor(ntile(value, 10))) %>%
  select(overall, metric, value, decile_rank, discipline) %>%
  ggplot(aes(x = decile_rank, y = overall)) +
    geom_boxplot() +
    stat_summary(fun.y=median, geom="line", aes(group=1), color = "red")  +
    stat_summary(fun.y=median, geom="point", aes(group=1), color = "red")  + 
    facet_grid(discipline~metric) +
    theme_minimal() +
    theme(
      strip.text = element_text(size = 12, face = "bold"),
      axis.title = element_text(size = 12)
    ) +
    labs(x = "Decile ranks",
         y = "Overall Teaching Quality")

ggsave("~/Desktop/aarmp_disc_boxplots.pdf", height = 8, width = 8)
```


Get the regression infomraiton
```{r}
l <- c("None", "Moderate", "High")

# This is the main formula we will be using, uses categorical research indicator variables
formula_discrete <- overall ~ 
            # Teacher Characteristics
            gender +
            scientific_age +
            mentions_accent +
            hotness +
            rank +
            # Course characteirstics
            difficulty +
            #difficulty_sd +
            interest +
            #interest_sd +
            #teaches_upper +
            mentions_ta +
            # Research characteristics
            citedness +
            all_output +
            grantiness +
            awardiness +
            # Disciplinary differences
            discipline + 
            # University
            uni_type +
            uni_control +
            review_count

df <- ready_data %>%
  mutate(
    rank = factor(rank, labels = c("Assistant", "Associate", "Full")),
    scientific_age = scale(scientific_age),
    uni_type = ifelse(is.na(uni_type) | uni_type != "r1", "Not R1", "R1"),
    discipline = factor(discipline, labels = c("Engineering", "Humanities", "Medical", "Natural", "Social")),
    difficulty = scale(difficulty),
    interest = scale(interest),
    citedness = factor(citedness, levels = l),
    all_output = factor(all_output, levels = l),
    output = factor(output, levels = l),
    bookiness = factor(bookiness, levels = l),
    awardiness = factor(awardiness, levels = l),
    grantiness = factor(grantiness, levels = l)
  ) %>%
  filter(review_count >= 5)

fit_research_discrete = lm(formula_discrete, data = df)

# used for copying numbers to text
broom::tidy(fit_research_discrete, conf.int = T) %>%
  mutate(estimate = round(estimate, 3),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2)) %>%
  select(term, estimate, conf.low, conf.high)


summary(fit_research_discrete)
```

```{r}
new_names = c(
  genderM = "Is Male",
  scientific_age = "Scientific Age",
  mentions_accentTRUE = "Mentions Accent = True",
  hotnesshot = "Has Chili Pepper",
  rankAssociate = "Rank = Associate",
  rankFull = "Rank = Full",
  difficulty = "Difficulty",
  interest = "Student Interest",
  mentions_taTRUE = "Mentions TA = True",
  citednessModerate = "Citedness = Moderate",
  citednessHigh = "Citedness = High",
  all_outputModerate = "Output = Moderate",
  all_outputHigh = "Output = High",
  grantinessModerate = "Grants Held = Moderate",
  grantinessHigh = "Grants Held = High",
  awardinessModerate = "Awards Won = Moderate",
  awardinessHigh = "Awards Won = High",
  disciplineHumanities = "Humanities",
  disciplineMedical = "Medical Sci.",
  disciplineNatural = "Natural Sci.",
  disciplineSocial = "Social Sci.",
  uni_typeR1 = "Uni. Type = R1",
  uni_controlpublic = "Uni. Control = Public",
  review_count = "#Reviews"
)


aarmp_reg_table <- stargazer(fit_research_discrete,
          covariate.labels = new_names,
          ci = TRUE,
          single.row = T,
          type = "latex", 
          style = "all2")

write(aarmp_reg_table, "~/Desktop/aarmp_reg_table.txt")

```

```{r}
ord <- names(fit_research_discrete$coefficients)[order(fit_research_discrete$coefficients, decreasing = T)]

summ <- summary(fit_research_discrete)


tidymodel <- broom::tidy(fit_research_discrete, conf.int = T)

p1 <- tidymodel %>%
  filter(term != "(Intercept)") %>%
  rowwise() %>%
  mutate(
    feature = ifelse(any(sapply(c("citedness", "awardiness", "grantiness", "all_output"), 
                          function(x) {return(grepl(x, as.character(term)))}
                          )
                      ), "Research", "Characteristic")
      ) %>%
  ungroup() %>%
  mutate(
    term = plyr::revalue(term, new_names),
    term = reorder(term, estimate)
  ) %>%
  ggplot(aes(x = term, y = estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  geom_hline(aes(yintercept = 0), linetype = "dotted", color = "black", size = 0.4) +
  ylim(-0.6, 0.6) +
  coord_flip() +
  scale_shape_manual(name = "", values = c(19, 1)) +
  scale_color_manual(name = "", values = c("black", "blue")) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    axis.title.y = element_blank(),
    legend.position = "none",
    #legend.position = c(0.75, 0.06),
    #legend.text = element_text(size = 12),
    #legend.background = element_blank(),
    #legend.box.background = element_rect(colour = "black")
    ) +
  labs(y = "Model coefficient")


ggsave("~/Desktop/aarmp_regression.pdf", p1, width = 4, height = 8)
p1
```


Get the regression infomraiton
```{r}
l <- c("None", "Moderate", "High")

# Here is the second fomula, this time using the continuous field-noramlized indicators
formula_continuous <- overall ~ 
            # Teacher Characteristics
            gender +
            scientific_age +
            mentions_accent +
            hotness +
            rank +
            # Course characteirstics
            difficulty +
            interest +
            mentions_ta +
            # Research characteristics
            norm_citations +
            norm_all_pubcount +
            norm_awardcount +
            norm_grantcount +
            # Disciplinary differences
            discipline + 
            # University
            uni_type +
            uni_control +
            review_count

df <- ready_data %>%
  mutate(
    rank = factor(rank, labels = c("Assistant", "Associate", "Full")),
    scientific_age = scale(scientific_age),
    uni_type = ifelse(is.na(uni_type) | uni_type != "r1", "Not R1", "R1"),
    discipline = factor(discipline, labels = c("Engineering", "Humanities", "Medical", "Natural", "Social")),
    difficulty = scale(difficulty),
    interest = scale(interest)
  ) %>%
  filter(review_count >= 5)

fit_research_continuous <- lm(formula_continuous, data = df)


# used for copying numbers to text
broom::tidy(fit_research_continuous, conf.int = T) %>%
  mutate(estimate = round(estimate, 3),
         conf.low = round(conf.low, 2),
         conf.high = round(conf.high, 2)) %>%
  select(term, estimate, conf.low, conf.high)


summary(fit_research_continuous)
```

```{r}
new_names = c(
  genderM = "Is Male",
  scientific_age = "Scientific Age",
  mentions_accentTRUE = "Mentions Accent = True",
  hotnesshot = "Has Chili Pepper",
  rankAssociate = "Rank = Associate",
  rankFull = "Rank = Full",
  difficulty = "Difficulty",
  interest = "Student Interest",
  mentions_taTRUE = "Mentions TA = True",
  norm_citations = "Citations",
  norm_all_pubcount = "Publications",
  norm_awardcount = "Awards",
  norm_grantcount = "Grants",
  disciplineHumanities = "Humanities",
  disciplineMedical = "Medical Sci.",
  disciplineNatural = "Natural Sci.",
  disciplineSocial = "Social Sci.",
  uni_typeR1 = "Uni. Type = R1",
  uni_controlpublic = "Uni. Control = Public",
  review_count = "#Reviews"
)


aarmp_reg_table_continuous <- stargazer(fit_research_continuous,
          covariate.labels = new_names,
          ci = TRUE,
          single.row = T,
          type = "latex", 
          style = "all2")

write(aarmp_reg_table_continuous, "~/Desktop/aarmp_reg_table_continuous.txt")

```


We can perform an ANOVA to compare the two models and find no real improvement in their ability to explain the data. 
```{r}
anova(fit_research_discrete, fit_research_continuous)
```

```{r}
ready_data_filt <- ready_data %>%
  gather(key, value, norm_all_pubcount, norm_citations, norm_dollars) %>%
  filter(value > 0)

ready_data_filt %>%
  ggplot(aes(x = value)) +
  stat_density() +
  coord_cartesian(xlim = c(0, 10)) +
  facet_wrap(~key) +
  geom_vline(aes(xintercept = quantile(value, 0.9)))
```

## Figure SI.9
```{r}
# get set of relevant variables
get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
}

# Select the variables that we are most interested in
selected <- ready_data[c(
  "overall", "norm_citations", "norm_articlecount", "norm_confproc", "norm_bookcount", "norm_awardcount", "norm_grantcount"
)]

# Setup the labels for the figure
labs <- c("RMP Quality", "Cites", "Articles", "Proceedings", "Books", "Awards", "Grants")

# Run the correlations (this will take a while)
correlations <- cor(selected, method="kendall", use="pairwise") 

# Get the triangle
tri_cor <- get_upper_tri(correlations)
mat_melt <- reshape2::melt(tri_cor, na.rm = T)
```



```{r}
# Plot the correlation matrix
p2 <- ggplot(data = mat_melt, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "black", size = 0.25) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", 
     midpoint = 0, limit = c(-1,1), space = "Lab", name = "Kendall\nCorrelation") +
  theme_bw() + 
  coord_fixed() +
  geom_text(aes(Var1, Var2, label = round(value, 3)), color = "black", size = 2.5) +
  theme(
    axis.text.x = element_text(angle = 45,  hjust = 0, size = 12, family = "Times"),
    axis.text.y = element_text(size = 12, family = "Times"),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_text(size = 12),
    #legend.title = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(1, 0.1),
    #legend.position = "left",
    #plot.margin = margin(10, 10, 10, 10),
    legend.direction = "horizontal"
    ) +
    guides(fill = guide_colorbar(barwidth = 5, barheight = 1,
           title.position = "top", title.hjust = 0.4)
    ) +      
  scale_x_discrete(labels = labs, position = "top") +
  scale_y_discrete(position = "left", expand = c(0, 0), labels = labs)

ggsave("~/Desktop/aarmp_kendall.pdf", p2, width = 5, height = 4)
p2
```




```{r}
# Load the tag data
tag_data <- read_csv("~/Desktop/AARMP_2017/processed_files/aarmp_tagdata.csv", 
    col_types = cols(X1 = col_skip()))


mat <- as.matrix(tag_data[2:21])
rownames(mat) <- tag_data$PersonId
melted <- reshape2::melt(mat)

melted <- melted[melted$value, ]
melted$Var2 <- tolower(melted$Var2)
overall_mat <- merge(melted, as.matrix(tag_data[c(22, 1)]), by.x = "Var1", by.y = "PersonId")

medians <- (overall_mat %>% 
  group_by(Var2) %>% 
  dplyr::summarize(v = median(overall)) %>%
  arrange(v))

overall_mat$Var2 <- factor(overall_mat$Var2, levels <- medians$Var2, ordered = T)

overall_mat <- overall_mat %>%
  inner_join(ready_data, by = c("Var1" = "id")) %>%
  dplyr::select(-scientific_age) # for some reason its really complaining about this variable, so I'm getting rid of it for now

p3 <- overall_mat %>%
  mutate(Overall = overall.x) %>%
  ggplot(aes(x = as.factor(Var2), y = Overall)) +
  geom_boxplot(outlier.alpha = 0.05,  alpha = 0.8, fill = "dodgerblue4") +
  theme_minimal() +
  labs(y = "Rating") +
  guides(fill = FALSE) +
  theme(
    axis.text = element_text(size = 12, family = "Times"),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 12),
    plot.title = element_text(margin = margin(t = 5, b = 5))
  ) +
  coord_flip()
  

ggsave("~/Desktop/aarmp_tags.pdf", p3, height = 5, width = 5)
p3
```

```{r}

plots <- list(p2, p3)
grobs <- list()
widths <- list()
heights <- list()

for (i in 1:length(plots)){
    grobs[[i]] <- ggplotGrob(plots[[i]])
    widths[[i]] <- grobs[[i]]$widths[2:5]
    heights[[i]] <- grobs[[i]]$heights[2:5]
}

maxwidth <- do.call(grid::unit.pmax, widths)
maxheight <- do.call(grid::unit.pmax, heights)

for (i in 1:length(grobs)){
     grobs[[i]]$widths[2:5] <- as.list(maxwidth)
     grobs[[i]]$heights[2:5] <- as.list(maxheight)
}

g1 <- do.call("grid.arrange", c(grobs, ncol = 1))

g <- grid.arrange(p1, g1, ncol = 2)
ggsave("~/Desktop/aarmp_figure.pdf", g, height = 8, width = 8)
```

```{r}
discipline_box <- ready_data %>%
  mutate(
    discipline = factor(discipline, 
                        levels = c("Engineering", "Natural Sciences", "Medical sciences", "Social Sciences", "Humanities"),
                        labels = c("Eng.", "Natural", "Medical", "Social", "Hum."))
  ) %>%
  ggplot(aes(x = discipline, y = overall)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Discipline",
    y = "Overall Teaching Quality"
    )
```


```{r}
chili_box <- ready_data %>%
  mutate(
    hotness = factor(hotness, labels = c("No Chili Pepper", "Has Chili Pepper"))
  ) %>%
  ggplot(aes(x = hotness, y = overall)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Has chili pepper",
    y = ""
    )
```


```{r}
gender_box <- ready_data %>%
  mutate(
    gender = factor(gender, labels = c("Female", "Male"))
  ) %>%
  ggplot(aes(x = gender, y = overall)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Gender",
    y = ""
    )
```


```{r}
uni_type_box <- ready_data %>%
  mutate(
    uni_type = ifelse(uni_type == "r1", "R1", "Not R1")
  ) %>%
  ggplot(aes(x = uni_type, y = overall)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Univeristy Type",
    y = "")
```


```{r}
uni_control_box <- ready_data %>%
  mutate(
    uni_control = factor(uni_control, labels = c("Private", "Public"))
  ) %>%
  ggplot(aes(x = uni_control, y = overall)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "University Control",
    y = "")
```


```{r}
rank_box <- ready_data %>%
  mutate(
    rank = factor(rank, labels = c("Assistant", "Associate", "Full"))
  ) %>%
  ggplot(aes(x = rank, y = overall)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold")
  ) +
  labs(
    title = "Professor rank",
    y = "")
```


```{r}
g <- grid.arrange(discipline_box, gender_box, chili_box, uni_type_box, uni_control_box, rank_box, ncol = 2)
ggsave("~/Desktop/aarmp_boxplots_fig.pdf", g, width = 8, height = 8)
```


```{r}
age_rank <- ready_data %>%
  ggplot(aes(x = rank, y = scientific_age)) +
  geom_boxplot() +
  theme_minimal() +
  theme(
    plot.title = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold")
  ) +
  labs(
    y = "Scientific Age")

ggsave("~/Desktop/aarmp_age_rank_boxplots.pdf", age_rank, width = 6, height = 5)
```